# Quality of Life Associations
Up to now we have explored how each quality of life indicator is associated with income inequality. It is equally important to disect how the quality of life indicators interplay with each other.

```{r}
fig<-data.melt.avg %>% dplyr::mutate_if(is.numeric, round, 2)%>%
  plotly::plot_ly(
             x=~education_discrepancy_avg,y=~babies_per_woman_avg,z=~suicide_avg,
             color=~continent,colors=region.color,alpha=.8,
             text=~country,
             hovertemplate=paste('%{x}','%{text}','GINI:%{y}','SR:%{z}',sep='<br>'))
fig <- fig %>% plotly::add_markers()
fig <- fig %>% plotly::layout(scene = list(xaxis = list(title = 'Education Discrepancy'),
                     yaxis = list(title = 'Babies per Woman'),
                     zaxis = list(title = 'Suicide Rate')))
fig <- fig %>% plotly::layout(title="Suicide Rate by GINI Index", showlegend = T)
fig
```

Figure \@ref(fig:violinplots) illustrates the distribution of each indicator by region. The colored-polygons represent the density for a given value. The box and whisker plot within each colored-polygon depicts the non-outlier quartile values. Note the different axes scales for each indicator.

```{r violinplots, fig.cap="Inequality to Quality of Life", echo=FALSE}
data.melt.comb <- data.melt %>% 
  dplyr::select(continent, gini, education_discrepancy, babies_per_woman, suicide) %>%
  reshape2::melt(id="continent")
  
plot.violin(data.melt.comb, highcharter::highchart(), "value", "continent", "variable", highcharter::data_to_boxplot(data.melt.comb, value, continent, variable), c(1, 0, 2, 3)) %>%
    highcharter::hc_yAxis_multiples(
      list(title = list(text = "Education Discrepancy"), maxRange=1),
      list(title = list(text = "Gini Coeficient")),
      list(title = list(text = "Babies"), opposite=T),
      list(title = list(text = "Suicide"), opposite=T))
```

Our analysis thus far has disregarded how the variables of interest and Gini coefficient change over time. The set of line plots below depict the time-related trends in each of the variables from 2000 to the present. 

```{r lineplots, fig.cap="Quality of Life over Time", echo = FALSE}

# Data grouped by year to be used for the plots
data <- data.melt.yearly

# Plot of education
fig.education <- data %>%
  plotly::group_by(continent) %>%
  plotly::plot_ly(x = ~year, color = ~continent, 
                  colors = c('#E41A1C', '#377EB8', '#4DAF4A', '#984EA3', "#FF7F00", "#FFFF33", "#A65628"),
                  legendgroup = ~continent, 
                  hovertext = paste("Education Discrepancy: ", data$education_discrepancy_avg, 
                                    "<br>Year: ", data$year, 
                                    "<br>Continent: ", data$continent),
                  hoverinfo = "text") %>%
  plotly::add_lines(y = ~education_discrepancy_avg, showlegend = FALSE) %>%
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = "Education Discrepancy"))

# Plot of suicide
fig.suicide <- data %>%
  plotly::group_by(continent) %>%
  plotly::plot_ly(x = ~year, color = ~continent, 
                  colors = c('#E41A1C', '#377EB8', '#4DAF4A', '#984EA3', "#FF7F00", "#FFFF33", "#A65628"),
                  legendgroup = ~continent, 
                  hovertext = paste("Suicide: ", data$suicide_avg, 
                                    "<br>Year: ", data$year, 
                                    "<br>Continent: ", data$continent),
                  hoverinfo = "text") %>%
  plotly::add_lines(y = ~suicide_avg) %>%
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = "Suicide"))

# Plot of birth rate
fig.babies <- data %>%
  plotly::group_by(continent) %>%
  plotly::plot_ly(x = ~year, color = ~continent, 
                  colors = c('#E41A1C', '#377EB8', '#4DAF4A', '#984EA3', "#FF7F00", "#FFFF33", "#A65628"), 
                  legendgroup = ~continent, 
                  hovertext = paste("Babies: ", data$babies_per_woman_avg, 
                                    "<br>Year: ", data$year, 
                                    "<br>Continent: ", data$continent),
                  hoverinfo = "text") %>%
  plotly::add_lines(y = ~babies_per_woman_avg, showlegend = FALSE) %>%
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = "Babies Per Woman"))

# Plot of Gini coefficient
fig.gini <- data %>%
  plotly::group_by(continent) %>%
  plotly::plot_ly(x = ~year, color = ~continent, 
                  colors = c('#E41A1C', '#377EB8', '#4DAF4A', '#984EA3', "#FF7F00", "#FFFF33", "#A65628"),      
                  legendgroup = ~continent, 
                  hovertext = paste("Gini Coefficient: ", data$gini_avg, 
                                    "<br>Year: ", data$year, 
                                    "<br>Continent: ", data$continent),
                  hoverinfo = "text") %>%
  plotly::add_lines(y = ~gini_avg, showlegend = FALSE) %>%
  plotly::layout(xaxis = list(title = "Year"), yaxis = list(title = "Gini Coefficient")) 

# Combine each of the four subplots into one plot
fig <- plotly::subplot(fig.education, fig.suicide, fig.babies, fig.gini, nrows=2, titleY = TRUE, margin = c(0.085,0.075,0.1,0.1)) %>% plotly::layout(title = "Income Inequality and Quality of Life over Time")

fig
```

Initially, we focused on only three continents: Africa, Asia, and Europe. (This view can be obtained by double-clicking on Africa and then single-clicking Asia and Europe.) For the most part, all four variables across the three continents exhibit a downward trend across time. Additionally, in the plots of education discrepancy, birth rate, and Gini coefficient, countries in Africa have the largest magnitudes, those in Europe have the smallest values, and those in Asia are in the middle. 

We then added North America and South America to our analysis. (This view can be obtained by single-clicking North America and South America.) Other than the suicide rate in South America, the four variables tend to exhibit a downward trend across time for all five continents. The primary observation, however, is how North America and South America have high income inequality relative to what is expected from their low birth rates and low levels of education inequality. 

Oceania and the Seven Seas consist of a small number of countries and, thus, contain noise that make it hard to identify time-related trends for education inequality. That said, the time-related trends for  birth rate and income inequality for these two regions are similar to those observed across the other five continents. 

Ultimately, we observe that all four variables across each continent have decreased over time, on average and with few exceptions. Additionally, the line plots extend our prior analysis to reveal that the associations between education inequality, birth rate, and income inequality have persisted over time. 
