---
output: html_document
editor_options: 
  chunk_output_type: console
---
# sdf

```{r}
library (dplyr)
library (ggplot2)
library (countrycode)
library (tidyverse)
library (ggpubr)
library (magrittr)
```

## Birth rates and contraception rates over time

Aggregate by year
```{r}
data.melt.avg.yearly <- data.melt %>%
  dplyr::select(-one_of("country", "subregion")) %>%
  dplyr::group_by(year, continent) %>% dplyr::summarise_each(list(avg = ~mean(.x, na.rm=T)))
```

Plot average birth rate and average gini coefficient by year across each continent
```{r}
# NOT WORKING
require(gridExtra)
plot1 <- ggplot(drop_na(data.melt.avg.yearly), aes(x=year, col = continent)) +
  geom_line(aes(y=babies_per_woman_avg))
plot2 <- ggplot(drop_na(data.melt.avg.yearly), aes(x=year, col = continent)) +
  geom_line(aes(y=gini_avg))

ggarrange(plot1, plot2, ncol=2, nrow=1, common.legend = TRUE, legend="bottom")
```

From this, we can see that the number of babies per woman and the amount of income inequality have, for the most part, decreased over time across all continents. 

## Scatter plots of birth rates and contraception rates by country
We're interested in investigating how contraception and birth rates are associated with income inequality (as measured by the gini coefficient).

A scatter plot demonstrating the relationship between contraception and income inequality. Each point represents a unique country. The size of the point represents the average income per person and the color corresponds to the continent the country is affiliated with. 

```{r}
g <- ggplot(drop_na(data.melt.avg), aes(x = contraception_avg, y = gini_avg, col = continent,  size = income_avg)) + geom_point(alpha = 0.7) #+ geom_smooth(method = lm, aes(group =1, colour = "black"))
plot(g)
```

The plot demonstrates a negative association between contraception use and income inequality. In other words, countries with higher contraception rates are those that have lower average gini coefficients. That said, we notice a similar association between contraception use and average income per person. So it is unclear which of the variables (income or income inequality) may explain this trend. 

Plotting the contraception against income we see a stronger trend than that observed plotting contraception against income inequality, which would indicate that income may be the primary driving factor explaining the observed trend. 
```{r}
g <- ggplot(drop_na(data.melt.avg), aes(x = contraception_avg, y = income_avg, col = continent,  size = gini_avg)) + geom_point() #+ geom_smooth(method = lm, aes(group =1, colour = "black"))
plot(g)
```





